Publicity Outreach for ``Quantum interval-valued probability: Contextuality and the Born rule''

### 2018-04-21 dissertation.tex
Keep updating...

%% LyX 2.2.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[twoside]{iuphd}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{verbatim}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[all]{xy}

@@

This marriage of quantum mechanics and computer science first envisioned
and popularized by Feynman has created an awkward, but opportune,
moment. The embarrassing dilemma was concisely described by Aaronson
with the following three statements \cite{AaronsonArkhipov2011,BFRDARW2013}:
\begin{enumerate}[label=(\roman{enumi})]
\item \label{enu:Textbook-quantum-mechanics}Textbook quantum mechanics
is correct. 
\item \label{enu:no-efficient-classical-factoring}There does not exist
an efficient classical factoring algorithm. 
\item \label{enu:extended-Church-Turing-Thesis}The extended Church-Turing
thesis — that probabilistic Turing machines can efficiently simulate
any physically realizable model of computation — is correct \cite{BernsteinVazirani1997,Kaye2007}. 
\end{enumerate}
There is overwhelming evidence to support each of these statements.
The theoretical framework of quantum mechanics \ref{enu:Textbook-quantum-mechanics}
has withstood decades of experimental confirmation. Entire industries
are founded on the assumption \ref{enu:no-efficient-classical-factoring}
that algorithms like RSA are secure and they also have withstood years
of attempted attacks \cite{boneh1999twenty,wiki:RSAFactoring}. Finally
the entire field of complexity theory in computer science which has
also withstood years of field testing rests, in essence, on assumption
\ref{enu:extended-Church-Turing-Thesis} \cite{Aaronson2005,Piccinini2015}.
And yet at least one of these three statements \emph{must be false}!
Indeed if there is a corresponding efficient classical factoring algorithm,
then we concede \ref{enu:no-efficient-classical-factoring}. If it
is correct Shor's efficient factoring algorithm is realizable, and
we can prove there is no efficient classical factoring algorithm,
we concede \ref{enu:extended-Church-Turing-Thesis}. Otherwise, if
we cannot implement Shor's algorithm no matter how hard we try, textbook
quantum mechanics, i.e., \ref{enu:Textbook-quantum-mechanics} may
need to be improved by a better theory. It is unlikely that there
will be a simple resolution to this awkward situation. It is more
likely that the resolution will emerge from deep and careful analyses
of the foundations of each field.

@@

For any vector $\ket{\Psi}=\begin{pmatrix}\alpha_{0} & \alpha_{1} & \ldots & \alpha_{D-1}\end{pmatrix}^{T}$,
the Hermitian dot product $\ip{\Psi}{\Psi}$ is equal to $\sum_{i=0}^{D-1}\fnorm{\alpha_{i}}$,
which is the sum of the field norms for the complex coefficients.
For convenience, we now extend the field norm to include vector arguments
by defining 
\begin{equation}
\fnorm{\ket{\Psi}}=\ip{\Psi}{\Psi}=\sum_{i=0}^{D-1}\fnorm{\alpha_{i}}\,.\label{fnormD.eq}
\end{equation}
The field norm of a vector can vanish for non-vanishing vectors.

>>>>

%% LyX 2.3.0 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[twoside]{iuphd}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{verbatim}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[all]{xy}

@@

This marriage of quantum mechanics and computer science first envisioned
and popularized by Feynman has created an awkward, but opportune,
moment. The embarrassing dilemma was concisely described by Aaronson
with the following three statements \cite{AaronsonArkhipov2011,BFRDARW2013}:
\begin{enumerate}[label=(\roman{enumi})]
\item \label{enu:Textbook-quantum-mechanics}Textbook quantum mechanics
is correct. 
\item \label{enu:no-efficient-classical-factoring}There does not exist
an efficient classical factoring algorithm. 
\item \label{enu:extended-Church-Turing-Thesis}The extended Church-Turing
thesis --- that probabilistic Turing machines can efficiently simulate
any physically realizable model of computation --- is correct \cite{BernsteinVazirani1997,Kaye2007}. 
\end{enumerate}
There is overwhelming evidence to support each of these statements.
The theoretical framework of quantum mechanics \ref{enu:Textbook-quantum-mechanics}
has withstood decades of experimental confirmation. Entire industries
are founded on the assumption \ref{enu:no-efficient-classical-factoring}
that algorithms like RSA are secure and they also have withstood years
of attempted attacks \cite{boneh1999twenty,wiki:RSAFactoring}. Finally
the entire field of complexity theory in computer science which has
also withstood years of field testing rests, in essence, on assumption
\ref{enu:extended-Church-Turing-Thesis} \cite{Aaronson2005,Piccinini2015}.
And yet at least one of these three statements \emph{must be false}!
Indeed if there is a corresponding efficient classical factoring algorithm,
then we concede \ref{enu:no-efficient-classical-factoring}. If it
is correct Shor's efficient factoring algorithm is realizable, and
we can prove there is no efficient classical factoring algorithm,
we concede \ref{enu:extended-Church-Turing-Thesis}. Otherwise, if
we cannot implement Shor's algorithm no matter how hard we try, textbook
quantum mechanics, i.e., \ref{enu:Textbook-quantum-mechanics} may
need to be improved by a better theory. It is unlikely that there
will be a simple resolution to this awkward situation. It is more
likely that the resolution will emerge from deep and careful analyses
of the foundations of each field.

@@

For any vector $\ket{\Psi}=\begin{pmatrix}\alpha_{0} & \alpha_{1} & \ldots & \alpha_{D-1}\end{pmatrix}^{T}$,
the Hermitian dot product $\ip{\Psi}{\Psi}$ is equal to $\sum_{i=0}^{D-1}\fnorm{\alpha_{i}}$,
which is the sum of the field norms for the complex coefficients~\cite{wiki:FieldNorm}.
For convenience, we now extend the field norm to include vector arguments
by defining 
\begin{equation}
\fnorm{\ket{\Psi}}=\ip{\Psi}{\Psi}=\sum_{i=0}^{D-1}\fnorm{\alpha_{i}}\,.\label{fnormD.eq}
\end{equation}
The field norm of a vector can vanish for non-vanishing vectors.

### 2018-04-21 revised_prop_letter_publicity.tex
Keep updating...

\maketitle
Randomness is exploited in cryptography, algorithmic complexity, and
gambling. While quantum physics is associated to randomness, it could
be causally deterministic if it were simulated by a classical non-contextual
hidden-variable model. The Kochen-Specker theorem excludes such models.
However, its main tenet has been questioned because finite-precision
measurements, i.e. errors in preparation or observation, might invalidate
its spirit. Our work, by introducing an interval-valued probabilistic
framework, establishes bounds on the validity of the Kochen-Specker
theorem in realistic, i.e., error-non-free, experimental environments.
The conclusion is that a quantum theory of measurement that includes
errors in the probability outcomes can still provide a valid test
of contextuality, as long as the error is bounded by a finite amount.

>>>>

\maketitle
Since randomness is exploited in cryptography, algorithmic complexity,
and gambling, people are always seeking reliable random sources. Quantum
physics could provide a reliable random source since it is widely
believed that the results of quantum measurements have no way to be
predicted deterministically. More precisely, this fact, called the
Kochen-Specker theorem, states that quantum physics cannot be simulated
by a classical non-contextual hidden-variable model. However, its
main tenet has been questioned because finite-precision measurements,
i.e., errors in preparation or observation, might invalidate its spirit.
Our work, by introducing an interval-valued probabilistic framework,
establishes bounds on the validity of the Kochen-Specker theorem in
realistic, i.e., error-non-free, experimental environments. The conclusion
is that a quantum theory of measurement that includes errors in the
probability outcomes can still be contextuality, as long as the error
is bounded by a finite amount. This also provides additional trust
of the randomness of quantum random sources.
